---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# cheedR 

<!-- badges: start -->
<!-- badges: end -->

A package for turning messy La Trobe University student enrolment extracts into tidy data for CHEEDR analysis.

## Introduction

One of the most painful parts of working with data at CHEEDR is extracting student data through the La Trobe University Business Intelligence (BI) interface. 

Bad news - this package cannot alleviate that pain. 

Good news - all the painful stuff after extracting data is much less horrid thanks to this package. Reading in extracted BI output, cleaning it, and joining external variables, such as secondary school data, and the Australian Bureau of Statistics (ABS) geodemographic variables, and producing a tidy finshed table are available within this package.   


## Installation

You can install a development version of `cheedR` from github with:

```{r, results='hide',message=FALSE}
#devtools::install_github("brcakitaki/cheedR")
#require(cheedR)
```

If you have been living under the proverbial rock and not yet installed the `tidyverse` suite of packages, do it, and then load it into the environment:

```{r, results='hide',message=FALSE}
# install.packages("tidyverse")
#require(tidyverse)
```

The cheedR package requires clean column names in the underscore_separated style, so we need the `janitor` package to do this:

```{r, results='hide', message=FALSE}
# install.packages("janitor")
#require(janitor)
```


We will use the nifty `skimr` package to skim over our data, so install and attach:

```{r, results='hide', message=FALSE}
# install.packages("skimr")
#require(skimr)
```

    

# 1. Getting the data

Lets walk through some of the routine steps for extracting student enrolment data at La Trobe: 

- Imagine you are analysing student retention by campus and over time, and you want to calculate 6 year completion rates. You want a comprehensive set of enrolment variables, over say 7 or 8 years. That is big data.   

- Open ghastly BI. First roadblock: a stingy size limit for data extraction. The size of the limit is something close to 250MB. You might spend a bit of time chasing this issue up with ICT, the vendor, and then your therapist. A couple of days, a teary phone call to mum, and 6 deep fried comfort meals later, you abandon fixing the root of problem. Gather your safety pins and duct tape, its time to find a work around.      

- One solution is to extract the data in batches -- a workflow this package is designed to accomodate. La Trobe University is organised into two colleges - the College of SHE and the College of ASCC - so I extract the data in batches by year and college. There is no strict rule here on what variables to split the batches on, although, because the package joins the enrolment tables by row, it is recommended that your batches all have the same columns. To extract 7 years of data, that is 14 extracts for the enrolment file. 

- Two variables require additional treatment before being joined to the enrolment data, so this workflow is designed to extract and read them in separately. The `postcode` variable and `parent_ed` variable require separate batches of extracts. 

- The `new_csv_dir` package creates directories to house this data. The `new_csv_dir` function sets up a folder in a path of your choosing (the default is your home directory) called 'Data_extracts_', with an added suffix for the date of the folder creation (this can be changed to any character you like using the `date_stamp` argument). Within this folder, the default function creates three additional folders: 
  1) 'Enrolments', where extracted enrolment tables should be saved.
  2) 'Parents', where extracted Parental Educational Attainment tables should be saved.
  3) 'Postcodes', where extracted Postcode  variables should be saved. 

- The functions allows you to choose whether you want to include a `postcode` or `parent` folder with a logical argument.

```{r}

#new_csv_dir(path = "~/", date_stamp = "1969", postcode = T, parent = T)

```

- Save the extractions into these folders as `.csv` files.                             



## Loading and joining the raw data

- The functions `enrolment_read`, `parent_read`, and `postcode_read`, will read in all the `.csv` files in the path you specify, and perform some basic cleaning processes. 

```{r}

#enrolment_df <- enrolment_read(path = "~/Data_extracts_1969/Enrolments")

#postcode_df <- postcode_read(path = "~/Data_extracts_1969/Postcodes")

#parents_df <- parent_read(path = "~/Data_extracts_1969/Parents")

```


```{r, include = FALSE}

# Load in toy data
#LTU.df <- read_csv("C:/Users/brcakitaki/Documents/LTU Enrolment data/Enrolment 18-20/LTU_2018_80000.csv") %>%
 # janitor::clean_names() %>% 
  #sample_frac(size = 0.5)

```


## Cleaning the data

Sometimes data extracted from BI contains cases that are just wrong. Other times the we may want to further simplify a variable. The variable 'latest_atar' often has several rogue ATARs in the mix, for instance some students are recorded with three or four digit long ATARs (i.e an ATAR of 5405). While there are some really smart kids in Australia, none as of yet have been able to get an ATAR larger than 100. So, we should trim the variable. The function 'ATAR_trim()', trims away ATAR values below 10, and above 100, and replaces them with NAs.

We often want an indicator of whether or not the student is Indigenous. The function 'Indigenous_var()' creates a new dichotomous Indigenous student variable from the BI variable 'aboriginal_or_torres_strait_islander_code'.  

The following code performs the cleaning: 

```{r example.2}

## Trim ATARs
#clean_df <- LTU.df %>% 
 # ATAR_trim() %>% 
  #Indigenous_var()

```

Lets see how our 'atar' variable has changed. In the output below, we see that the 'atar' variable has trimmed ATARs below 10 and above 100.

```{r}
#clean_df %>%
 # select(latest_atar,
  #       atar) %>% 
  #skim()
```

Now lets have a look at our new 'Indigenous' variable. The 'Indigenous' variable is now dichotomous, compared to the previous 5 value categorical.  

```{r}
#clean_df %>% 
 # select(aboriginal_or_torres_strait_islander_code,
  #       Indigenous) %>% 
  #table()

```


## Joining external variables


## Seeing the data






